High-level cognition is supported by information-rich but compressible brain activity patterns

### 2023-12-19 paper/main.tex
updated stats (correct for multiple comparisons) in decoding_and_compression notebook   - note: still need to update stats in paper

\title{High-order cognition is supported by information-rich but compressible brain activity patterns}

@@

  \caption{\textbf{Information content and compressibility.} \textbf{A.
  Variance explained for two images.} We applied principal components analysis
  to a photograph and drawing, treating the rows of the images as
  ``observations.'' Across columns, we identified the numbers of components
  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
  each image (the 100\% columns denote the original images). The numbers of
  components are indicated in parentheses, and the resulting ``compressed''
  images are displayed. \textbf{B. Representing two images with different
  numbers of components.} Using the same principal component decompositions as
  in Panel A, we computed the cumulative proportion of variance explained with
  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
  explained versus number of components.} For the images displayed in Panels A
  and B, we plot the cumulative proportion of variance explained as a function
  of the number of components used to represent each image. \textbf{D.
  Information rate and compressibility.} Across multiple images, the
  information rate (i.e., the amount of information contained in each image;
  horizontal axis) is high when each individual pixel provides information that
  cannot be inferred from other pixels. High-information rate images tend to be
  high-resolution, and low-information rate images tend to be low-resolution.
  Compressibility is related to the difference between the information required
  to specify the original versus compressed images (vertical axis). Highly
  compressible images often contain predictable structure (redundancies) that
  can be leveraged to represent the images much more efficiently than in their
  original feature spaces.}

\label{fig:information-compression} 
\end{figure}

@@

>>>>

\title{High-level cognition is supported by information-rich but compressible brain activity patterns}

@@

  \caption{\textbf{Information content and compressibility.} \textbf{A.
  Variance explained for two images.} Consider two example images: a photograph
  and simple drawing. The images have the same resolutions, but their content
  is very different. The photograph is more information rich: when we compare
  the original photograph and drawing (leftmost column), we can see that the
  photograph captures much more detail. We can also apply principal components
  analysis to the images, treating the rows of the images as ``observations.''
  Across columns, we identified the numbers of components required to explain
  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the
  100\% columns denote the original images). The numbers of components are
  indicated in parentheses, and the resulting ``compressed'' images are
  displayed. This analysis reveals that the drawing is more compressible: just
  16 components can explain 95\% of the variance in teh drawing, whereas 50
  components are required to explain 95\% of the variance in the photograph.
  \textbf{B. Representing two images with different numbers of components.}
  Using the same principal component decompositions as in Panel A, we computed
  the cumulative proportion of variance explained with 380 (original images),
  20, 4, or 2 components. This analysis provides another way of characterizing
  the compressibility of each image by showing that the same number of
  components can explain more variance in the drawing than in the photograph.
  \textbf{C. Template data from four synthetic datasets.} We constructed four
  synthetic datasets, each comprising 25 features (columns) observed across 100
  samples (rows) from each of 10 simulated ``participants.'' The datasets were
  constructed to contain different levels of informativeness and
  compressibility (see \textit{Synthetic data}). \textbf{D. Decoding accuracy
  by number of components.} For each synthetic dataset, we trained
  across-participant classifiers to decode timepoint labels.  Each panel displays the
  decoding accuracy as a function of the number of components used to represent
  the data. Error ribbons denote bootstrap-estimated 95\% confidence intervals.}
\label{fig:information-compression} 
\end{figure}

@@

\subsection*{Synthetic data}

To help illustrate the relationship between informativeness and compressibility
(Fig.~\ref{fig:information-compression}), we generated four synthetic datasets,
varying in informativeness and compressibility. Each dataset comprised
simulated observations of $K = 25$ features across $N = 100$ timepoints, from each of $S = 10$
participants. To create each dataset, we first constructed a ``template''
matrix of $N$ timepoints by $K$ features. We then generated participant-specific
data by adding independent noise to each entry in template matrix, drawn from
the unit normal distribution (i.e., with a mean of 0 and a variance of 1). We
repeated this process for each participant, yielding $S$ participant-specific
matrices for each dataset.

Since we estimate informativeness using the temporal decoding accuracy across
participants, highly informative data will tend to have observations that are
highly timepoint specific. Relatively uninformative data, in contrast, will
tend to have more similar observations across timepoints. To generate data with
``high informativeness,'' we constructed template matrices whose rows
(observations) were drawn independently from zero-mean multivariate normal
distributions. The covariances of these distributions were determined according
to the desired compressibility of the data, as described below. We used a
multi-step process to generate data with ``low informativeness.'' First we
generated new template matrices using the same procedure as for the ``high
informativeness'' datasets. We then multiplied each matrix by a constant ($\rho
= 0.1$) and computed the cumulative sum of each matrix's rows. This yielded
matrices whose rows were highly similar across observations.

Compressibility reflects the extent to which decoding accuracy is affected by
reducing the number of components used to represent the data. Highly
compressible data will tend to exhibit more similarities across features,
whereas less compressible data will tend to show greater independence across
features. To generate data with ``high compressibility,'' we set the covariance
matrix of the multivariate normal distribution to a toeplitz matrix whose first
row was given by $\left[K, K - 1, ..., 1\right]$. To generate data with ``low
compressibility,'' we set the covariance matrix to the identity matrix.

Template matrices for datasets with high informativeness and high
compressibility, high informativeness and low compressibility, low
informativeness and high compressibility, and low informativeness and low
compressibility are displayed in Figure~\ref{fig:information-compression}C. The
corresponding decoding curves are displayed in
Figure~\ref{fig:information-compression}D.
